Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-04-22 text/back_end_KS.qmd
LA rq1

The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in Figure X by total population across the three census periods. The Los Angeles region is one of the only multi-nuclei cases metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982. As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie. These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft. Worth Metroplex).

>>>>

The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods. The Los Angeles metropolitan area is one of the only multi-nuclei models in the country that is formed when urban growth occurs around multiple centers that serve different functions rather than expanding from a single central business district core (CITE- Harris and Ullman 1945). As depicted in @fig-la-maps, Geomarkets closely follow this overall metropolitan model. Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring of made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets. The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland.

The population growth in the Los Angeles metropolitan area increased from approximately 8 Million in 1980 to more than 12 Million in 2020. @fig-dallas-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model. The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernadino, and Ontario. The South and South Central Los Angeles Geomarket grew from approximately 1.2 to 1.6 Million from 1980 to 2020. In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets grew from approximately 700,000 to 1.2 Million and from 1.1 Million to nearly 3 Million during this time, respectively.
